\begin{center} \bf \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm] \end{center}

### 2021-05-03 main.tex
draft last sections

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{standalone}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{hyperref}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}

@@

\tikzset{
  treenode/.style = {align=center, inner sep=0pt, text centered,
    font=\sffamily},
  arn_n/.style = {treenode, circle, white, font=\sffamily\bfseries, draw=black, inner sep=-6pt,
    fill=black, text width=1.5em},% arbre rouge noir, noeud noir
  arn_r/.style = {treenode, circle, red, fill=red, 
    text width=1.5em, very thick, inner sep=-6pt},% arbre rouge noir, noeud rouge
  arn_x/.style = {treenode, rectangle, draw=black,
    minimum width=0.5em, minimum height=0.5em}% arbre rouge noir, nil
}

@@

@@

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.65\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
  \end{subfigure}%
  \begin{subfigure}{.35\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
  \end{subfigure}
\caption{{\bf The evolution of cooperation for different benefit values.} 
Here, we vary the benefit of defection $b$. In all cases, stochastic payoff
evaluation tends to reduce the evolving cooperation rates. Unless explicitly
varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$,
$\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$
time steps for each parameter
combination.}\label{fig:cooperation_rate_over_benefit}
\end{figure}

@@

In the game stage, each player is randomly matched with some other player in
the population to interact in a number of instances of the game
\(\begin{pmatrix} R & S  \\ T & P \end{pmatrix}\). The number of interactions is
not fixed, on the contrast we consider that another interaction has a
probability $\delta$ to continue following each turn. We assume herein that
individuals at most make use of simple \textit{reactive strategies} make
decisions in each round. Reactive strategy are a set of memory-one strategies
that only take into account the previous action of the opponent. Reactive
strategies can be written explicitly as a vector \(\in \R_{3}\). More
specifically a reactive strategy \(s\) is given by \(s=(y, p, q)\) where \(y\)
is the probability that the strategy opens with a cooperation and \(p, q\) are
the probabilities that the strategy cooperates given that the opponent
cooperated and defected equivalently.

In the updating stage, two players are randomly drawn from the population, a
`learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, we
assume the learner adopts the role model's strategy with probability

\begin{equation} \label{Eq:rho}
\rho(u_L, u_{RL}) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_{RL}) \big]}.
\end{equation}

where $\beta\!\ge\!0$ corresponds  to relative influence of the payoffs on the
adopt the strategy of the other is controlled by an external parameter, the so
called intensity of selection.

We iterate this basic evolutionary step until either the mutant strategy goes
extinct, or until it fixes in the population (in which case the mutant strategy
becomes the new resident strategy). After either outcome, we introduce a new
mutant strategy (uniformly chosen from all reactive
strategies at random), and we set the number of mutants to $k\!=\!1$. This
process of mutation and fixation/extinction is then iterated many times.xw The
fixation probability of the mutant strategy then takes the standard form
\cite{nowak2004emergence},

\begin{equation}
\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
\end{equation}

@@

We compare this process for stochastic payoff evaluation with the analogous
process where players update their strategies with respect to their {\it
expected} payoffs.

\subsection{Expected Payoffs}

Consider two players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
$s_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
continuation probability $\delta$, the probability that are in each of the
four possible states in the last round is given by the stationary distribution
\(\langle\mathbf{v}\) of the transition matrix \(M\).

@@

The long run steady state probability vector \(\langle\mathbf{v}\) is the solution
to \(\langle\mathbf{v}M=\langle\mathbf{v}\), can be combined with the payoff
\(u\) to give the payoffs for each player. More specifically,

\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\]

is the payoff that \(s_1\) achieves when interacting with \(s_2\). In
the evolutionary process given \(k\) individuals with mutant strategy \(s_M\)
and \(N - k\) individuals with the resident strategy \(s_R\) the expected
payoffs are,

\begin{equation} \label{Eq:ExpPay}
  \begin{array}{lcrcr}
  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
  \end{array}
\end{equation}

In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
expected payoffs has been considered in~\cite{imhof2010stochastic}.

\subsection{Stochastic Payoffs}

Initially, we consider the case that an individual's fitness is based on their
last interaction with one other individual. There
only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
Consider two players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
$s_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
continuation probability $\delta$, the probability that are in each of the
four possible states in the last round is given by:

\begin{equation}
    \mathbf{v}(s_1,s_2)\!=\!\Big(\mathbf{v}_{R}(s_1,s_2),\mathbf{v}_{S}(s_1,s_2),\mathbf{v}_{T}(s_1,s_2),\mathbf{v}_{P}(s_1,s_2)\Big).
\end{equation}

where,

@@

\begin{proof}

Assume a repeated prisoner's dilemma between two reactive strategies. Given the
continuation probability $\delta$, probability that the game ends in the after
the first round $(1 - \delta)$ and the expected distribution of the four
outcomes in the very first round is $\mathbf{v_0}$ defined as. Following the
first round the, the outcome of the next rounds with a probability $\delta$
is \(M\) such that,

\[\dots\]

It can shown that, \((1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.

\end{proof}

Equation~\ref{Eq:LastRound} can be generalised to all cases where we consider
the last \()

Equation~\ref{Eq:LastRound} is the probability vector that players \(s_1, s_2\)
are in each of the possible states of the last round. Given the population \(N\)
with \(k\) mutants in the last round there only possible combinations of interactions
are:

\begin{figure}
    \centering
    \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
\end{figure}

\subsection{Fixation probabilities under stochastic payoff evaluation}
Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
\begin{equation}
\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
\end{equation}
In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant. The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds. We use $x(u_1,u_2)$ to denote the probability that the randomly chosen resident obtained a payoff of $u_1$ in the last round of his respective game, and that the mutant obtained a payoff of $u_2$. Given that the payoffs are $u_1$ and $u_2$, the imitation probability is then given by $\rho(u_1,u_2)$, as specified by Eq.~(\ref{Eq:rho}). The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
\begin{equation}
\setlength{\arraycolsep}{1pt}
\begin{array}{llrl}
x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
&+	
&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
 \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
 \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
\end{array}
\end{equation}

The first term on the right side corresponds to the case that the learner and
the role model happened to be matched during the game stage, which happens with
probability $1/(N\!-\!1)$. In that case, we note that only those payoff pairs
can occur that are feasible in a direct interaction, $(u_1,u_2)\in
\mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by
the respective indicator function. Otherwise, if the learner and the role model
did not interact directly, we need to distinguish four different cases,
depending on whether the learner was matched with a resident or a mutant, and
depending on whether the role model was matched with a resident or a mutant.

>>>>

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{standalone}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[ruled,vlined]{algorithm2e}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}

@@

\tikzset{
  treenode/.style = {align=center, inner sep=0pt, text centered,
    font=\sffamily},
  arn_n/.style = {treenode, circle, white, font=\sffamily\bfseries, draw=black, inner sep=-6pt,
    fill=black, text width=1.5em},% arbre rouge noir, noeud noir
  arn_r/.style = {treenode, circle, red, 
    text width=1.5em, very thick, inner sep=4pt},% arbre rouge noir, noeud rouge
  arn_x/.style = {treenode, rectangle, draw=black,
    minimum width=0.5em, minimum height=0.5em}% arbre rouge noir, nil
}

@@

\newcommand{\splitatcommas}[1]{%
  \begingroup
  \begingroup\lccode`~=`, \lowercase{\endgroup
    \edef~{\mathchar\the\mathcode`, \penalty0 \noexpand\hspace{0pt plus 1em}}%
  }\mathcode`,="8000 #1%
  \endgroup
}

@@

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.65\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
  \end{subfigure}%
  \begin{subfigure}{.35\textwidth}
    \vspace{.3cm}
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
  \end{subfigure}
\caption{{\bf The evolution of cooperation for different benefit values.} 
Here, we vary the benefit of defection $b$. In all cases, stochastic payoff
evaluation tends to reduce the evolving cooperation rates. Unless explicitly
varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$,
$\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$
time steps for each parameter
combination.}\label{fig:cooperation_rate_over_benefit}
\end{figure}

@@

We consider a pairwise comparison process. We assume a population of \(N\)
individuals (\(N\) is even) where at any point in time there are at most two
different strategies in present. There are \(k\) individuals who play the mutant
strategy and \(N - k\) individuals who play the resident strategy. Each step of
the evolutionary process consists of two stages; a game stage and an update
stage.

In the game stage, each individual is randomly matched with some other
individual in the population. Their interaction lasts for a number of turns
which is not fixed but depends on the continuation probability \(\delta\). At
each turn the individuals have choose between cooperation (C) and defection (D).
If both players cooperate they receive the reward payoff \(R\), whereas if both
players defect they receive the punishment payoff \(P\). If one cooperates but
the other defects, the defector receives the temptation to defect, \(T\),
whereas the cooperator receives the sucker's payoff, \(S\).

In this work we consider different social dilemmas represented by the harmony
game, the stag hunt game, the snowdrift or hawk-dove game, the prisoner's
dilemma, and a special case of the prisoner's dilemma, the donation game
(Table~\ref{table:social_dilemmas}).

\begin{table}[!htbp]
  \begin{center}
    \resizebox{.65\textwidth}{!}{
    \renewcommand{\arraystretch}{1.5}
      \begin{tabular}{ccc}
  \toprule
  & \textbf{social dilemmas} & \textbf{payoffs' constrains} \\
  \midrule
(i) & stage hunt  & \(R > T > P > S\) \\
(ii) & snowdrift  & \(T > R > S > P\) \\
(iii) & harmony  & \(R > T > S > P\) \\
(iv) & prisoner dilemma  & \(T > R > P > S\) \\
(v)& donation game  & \(T > R > P > S\) \& \(T=b, R=b-c, S=-c, P=0\) \\
  \bottomrule
      \end{tabular}}
  \end{center}
  \caption{We present results on the all of the listed social dilemmas. The payoffs'
  constrains are different for each game.}
  \label{table:social_dilemmas}
\end{table}

Results for cases (i) - (iv) were presented in section~\ref{section:2_by_2_games}
and results for case (v) were presented in section~\ref{section:donation}.

A further assumption of our model is that individuals make use of reactive
strategies when they make decisions in each round. Reactive strategy are a set
of strategies that take into account only the previous action of the opponent.
A reactive strategy can be written explicitly as a vector,

\[s=(y, p, q)\]

where \(y\) is the probability that the strategy opens with a cooperation and
\(p, q\) are the probabilities that the strategy cooperates given that the
opponent cooperated and defected equivalently.

In the updating stage, two players are randomly drawn from the population, a
`learner' and a `role model'. The learner adopts the role model's strategy
based on the Fermi distribution function, %ToDo add reference

\begin{equation} \label{Eq:rho}
\rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
\end{equation}

where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the intensity of
selection.

We iterate this basic evolutionary step until either the mutant strategy goes
extinct, or until it fixes in the population and becomes the new resident
strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
strategy which is uniformly chosen from all reactive strategies at random.

This process of mutation and fixation/extinction is iterated many times.
Instead of simulating each step of the evolutionary process, we estimate
the probability that a newly introduced mutant fixes~\cite{nowak2004emergence}.
This is defined as the fixation probability of the mutant, and the standard
form is the following,

\begin{equation}\label{eq:appendix_fixation_probability}
\varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
\end{equation}

@@

This evolutionary process is summarized by Algorithm~\ref{algorithm:pairwise_comparison}.

\begin{algorithm}[!htbp]
  \SetAlgoLined
   $N \leftarrow$ population size\;
   $k \leftarrow 1$\;
   resident $\leftarrow (0, 0, 0)$\;
   \While{step $<$ maximum number of steps}{
    mutant $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\;
    fixation probability $\leftarrow \varphi $\;
    \If{$\varphi >$ random: $i \rightarrow [0, 1]$}{
      resident $\leftarrow$ mutant;
     }}
     \caption{Pairwise comparison process}\label{algorithm:pairwise_comparison}
\end{algorithm}

The aim of this work is to explore the effect of updating memory on the
cooperation rate of the evolved population. For this reason we consider two
different approaches when estimating the payoffs at the updating stage. The two
approaches we consider are those of (i) the expected payoffs and (ii) the
stochastic payoffs.

\subsection*{Expected Payoffs}

The expected payoffs are the conventional payoffs used in the updating
stage~\cite{imhof2010stochastic}. They are defined as the mean payoff an
individual in a well-mixed population that engages in repeated games with all
other population members.

We first define the payoff of two reactive strategies at the game stage. Assume
two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
It is not necessary to simulate the play move by move. Instead the play between
the two strategies is defined a Markov matrix \(M\),

@@

whose stationary vector \(\mathbf{v}\), combined with the payoff \(U\), yields
the expected outcome for each strategy,
\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\).

% In each step
Given a population of \(N\) individual with \(k\) individuals playing the mutant
strategy, \(s_M = (y_M, p_M, q_M)\), and \(N -k\) playing the resident strategy,
\(s_R = (y_R, p_R, q_R)\), the expected payoffs of a resident \(\pi_R\) and of
a mutant strategy \(\pi_M\) are defined as,

\begin{equation} \label{Eq:ExpPay}
  \begin{array}{lcrcr}
  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{U}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{U}\rangle,\\[0.5cm]
  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{U}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{U}\rangle.\\
  \end{array}
\end{equation}

Note that \(N\!-\!k\!-\!1\) is the number of other residents in the population,
and \(k\) is the number of residents. The payoffs are averaged by the number of
individuals in the population, given that interactions are not possible.
The same applies for the payoff of a mutant.

Given the expected payoff, the probabilities that the number of mutants
decreases and increases  (Equation~(\ref{eq:appendix_fixation_probability})),
are given by:

\begin{align*} 
  \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
  \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
\end{align*}

\subsection*{Stochastic Payoffs}

We compare the expected payoffs with the finite memory payoffs which we refer to
as the stochastic payoffs. Initially, we consider the case of where the payoffs
correspond to the last round payoff an individual achieved after interacting
with one other individual.

The probability of being in any of the outcomes \(R, S, T, P\) in the last round
is given by Equation~(\ref{Eq:LastRound})
(Proposition~\ref{proposition:last_round}).

\begin{Prop}\label{proposition:last_round}
Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
continuation probability \(\delta\). Then the probability of being in any of the
four outcomes \(R, S, T, P\) are given by,

@@

In these expressions, we have used
the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$,
and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
\end{Prop}

\begin{proof}
Given a play between two reactive strategies with continuation probability
$\delta$. The probability that the game ends on round \(t\) is given by,

\begin{equation}\label{eq:}
  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)}.
\end{equation}

Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
the very first round, and \(1- \delta\) the probability that the game ends.
It can be shown that,

\begin{align*}
  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
   & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
   & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
\end{align*}

and with some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
\end{proof}

Considering that individuals can only interact with one other individual,
at that in each step a mutant and a resident are selected there are only five
possible pairings:

\begin{itemize}
  \item the selected resident is paired with the selected mutant
  \item the selected resident is paired with a resident and the selected mutant is paired with a resident
  \item the selected resident is paired with a resident and the selected mutant is paired with a mutant
  \item the selected resident is paired with a mutant and the selected mutant is paired with a resident
  \item the selected resident is paired with a mutant and the selected mutant is paired with a mutant
\end{itemize}

Each of the pairings happens with a given probability and that is illustrated
by Figure~\ref{fig:single_pairs}.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
  \caption{The tree}\label{fig:single_pairs}
\end{figure}

The probability that the respective payoffs of the players
are given by $u_1$ and $u_2$ can be calculated as

\begin{equation}\label{eq:Chi}
\setlength{\arraycolsep}{1pt}
\begin{array}{llrl}
x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
&+	
&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
 \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
 \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
\end{array}
\end{equation}

The first term on the right side corresponds to the case that the learner and
the role model happened to be matched during the game stage, which happens with
probability $1/(N\!-\!1)$.In that case, we note that only those payoff pairs
can occur that are feasible in a direct interaction, $\splitatcommas{(u_1,u_2)\in
\mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}}$, as represented by
the respective indicator function. Otherwise, if the learner and the role model
did not interact directly, we need to distinguish four different cases,
depending on whether the learner was matched with a resident or a mutant, and
depending on whether the role model was matched with a resident or a mutant.

Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$
and that the remaining $k$ players use the mutant strategy
$S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by
one in one step of the evolutionary process can be written as

\begin{align}
\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2), \\
\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
\end{align}

In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen
learner is a resident, and $k/N$ is the probability that the role model is a
mutant. The sum corresponds to the total probability that the learner adopts the
role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two
player may have received in their respective last rounds. We use $x(u_1,u_2)$ to
denote the probability that the randomly chosen resident obtained a payoff of
$u_1$ in the last round of his respective game, and that the mutant obtained a
payoff of $u_2$.

This framework can be extended to consider the case of where the payoffs
correspond to the last \(n\) rounds payoff an individual achieved after
interacting with \(m\) other individuals. For the case \(n=2\) the payoffs
at the game stage are,

\begin{Prop}\label{proposition:last_two_rounds}
Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
continuation probability \(\delta\). Then the probability of being in any of the
sixteen outcomes \(\splitatcommas{RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR}\)
on the last two rounds are given by,

\begin{equation}
  \mathbf{v_{a_1, a_2}} = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad \text{ for } m_{a_1, a_2} \in M \ \& \ a_1, a_2 \in \{R, S, T, P\}
\end{equation}

\end{Prop}

Proposition~\ref{proposition:last_two_rounds} can be extended to the last \(n\)
rounds.

\begin{Prop}\label{proposition:last_n_rounds}
Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
continuation probability \(\delta\). Then the probability of being in any of the
sixteen outcomes \(\splitatcommas{RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR}\)
on the last two rounds are given by,

\begin{equation}
  \mathbf{v_{a_1, a_2}} = (1 - \delta) \prod m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}
\end{equation}

for \(m_{a_1, a_2} \in M\) and \(a_1, a_2 \in [1, 4]\).
\end{Prop}

Equation~\ref{eq:Chi} can also be extended to include interactions with two other
individuals. The possible pairings are illustrated by Figure~\ref{}.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.65\textwidth]{static/tree.tex}
  \caption{The tree}\label{fig:two_pairs}
\end{figure}

\section{Verifying analytical results with simulations}

The analytical results presented in this work have been verified with simulations.
More specifically the probabilities of Equation~(\ref{Eq:LastRound}),

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.75\textwidth]{static/stationary_four.pdf}
\caption{}
\end{figure}

Proposition~\ref{proposition:last_two_rounds},

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/stationary_sixteen.pdf}
\caption{}
\end{figure}

% \begin{Prop}\label{proposition:last_n_rounds}

% \end{Prop}
